ECE 612, Information Theory

### 2024-04-09 information-theory/main.tex
Draft of homework 5
differential entropy

@@

\section{Differential entropy}
\begin{theorem}[Differential entropy of Gaussian distribution]
    Let $X$ be Gaussian $N(0, \sigma^2)$, then

    $$
    h(X) = \frac{1}{2}\log{(2\pi e \sigma^2)}
    $$
\end{theorem}

\begin{theorem}
    Let $\mathbf{X}$ follow joint Gaussian distribution $N(\mathbf{0}, K)$, then:

    \begin{equation*}
        h(\mathbf{X}) = \frac{1}{2}\log ((2\pi e)^n \det{K})
    \end{equation*}
\end{theorem}

@@

\section{Rate distortion theory}
\end{document}

>>>>

\begin{theorem}
    Let $X, Y$ be random variables. $I(X; Y)$ is concave with respect to the probability distribution of X. For a fixed marignal distribution of $X$, $I(X; Y)$ is convex with respect to $f_{Y \mid X}$
\end{theorem}

@@

\section{Differential entropy}
\begin{definition}
    Let $X \in \mathbb{R}$ be a random variable with probability density function $f_X$. The \textbf{differential entropy} of $X$ is defined by

    \begin{equation*}
        h(X) = -\int_{x \in \mathbb{R}} f_X(x)\ln(f_X(x))dx = -E[\ln(f_X(X))]
    \end{equation*}
\end{definition}

Some notable results:

\begin{itemize}
    \item For uniform distribution over $[0, a]$, $f_X(x) = \frac{1}{a}$, $h(X) = \ln(a)$
    \item For $X \sim N(0, \sigma^2)$, $h(X) = \frac{1}{2}\ln(2\pi e \sigma^2)$
    \item For multivariate normal $\mathbf{X} \sim N(\mathbf{0}, K)$, $h(\mathbf{X}) = \frac{1}{2}\log ((2\pi e)^n \det{K})$
    \item For some constant $a$, $h(X + a) = h(X)$
    \item For some constant $a$, $h(aX) = \ln(\norm{a}) + h(X)$
\end{itemize}

\begin{theorem}
    Let $X$ be continuous random variable with $0$ mean and $\sigma^2$ variance, then

    \begin{equation*}
        h(X) \leq \frac{1}{2}\ln(2\pi e \sigma^2)
    \end{equation*}

    Equality is reached if and only if $X$ is Gaussian
\end{theorem}

@@

\section{Rate distortion theory}
\begin{theorem}
    Let $X$ follow $\operatorname{Bernoulli}(p)$, then

    $$
    R(D) = \begin{cases}
        h(p) - h(D)  & \text{When $D < \min(p, 1-p)$} \\
        0   &  \text{otherwise}
    \end{cases}
$$
\end{theorem}

\begin{theorem}
    For $X \leftsample N(0, \sigma^2)$:

    $$
    R(D) = \begin{cases}
        \frac{1}{2}\log(\frac{\sigma^2}{D}) & (D \leq \sigma^2) \\
        0 & \text{otherwise}
    \end{cases}
    $$
\end{theorem}
\end{document}
